---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Main Effects of Exam Stress on Vocal Production

## F0

Between-person variability in baseline F0 was substantial (τ₁ = 19.86281 Hz, 95% CI 
[17.68257, 22.44086]), accounting for 83% of total variance (ICC = 0.83). Individual 
differences in stress reactivity were moderate (τ₂ = 1.080926 Hz, 95% CI [0.04109089, 4.452379]), 
with a τ₂/β₁ ratio of 0.33, indicating that individual variability was approximately 
one-third the magnitude of the average stress effect. Importantly, using the posterior 
distribution of random slopes, we estimated that 95% of individuals exhibited 
stress-related F0 increases ranging from 1.2 to 5.4 Hz, indicating that while the 
magnitude varied across individuals, the direction of the effect was consistent: 
all participants showed pitch elevation under stress.

## NNE

Between-person variability in baseline NNE was substantial (τ₁ = [2.146829] dB, 95% CI 
[1.840238, 2.508342]), accounting for 55% of total variance (ICC = 0.55). Critically, 
individual differences in stress reactivity were very large (τ₂ = [0.8301103] dB, 95% CI 
[0.0432476, 1.911727]), with a τ₂/β₁ ratio of 1.27, indicating that individual heterogeneity exceeded the magnitude of the average population effect. 

Using the posterior distribution of random slopes, we estimated that 95% of individuals exhibited stress-related NNE changes ranging from -2.3 to +1.0 dB. This wide prediction interval crosses zero, revealing substantial heterogeneity in individual responses. 
Specifically, we estimate that approximately 22% of individuals showed stress-related 
increases in glottal noise (positive NNE changes), contrary to the population-level 
trend toward noise reduction. This pronounced between-person heterogeneity contrasts 
sharply with the consistency of F0 responses, where all individuals showed pitch 
elevation under stress, and suggests that phonatory control adjustments are more 
strongly influenced by individual differences than arousal-driven pitch changes.


| Metric | F0 Mean | NNE |
|--------|---------|-----|
| Population effect (β₁) | +3.3 Hz*** | -0.65 dB*** |
| Between-person SD (τ₂) | 1.1 Hz | 0.83 dB |
| Ratio τ₂/β₁ | 0.33 | 1.27 |
| ICC (baseline) | 82.6% | 54.5% |
| 95% PI for individual effect | [1.2, 5.4] Hz | [-2.3, +1.0] dB |
| % with opposite effect | 0% | 21.6% |
| **Interpretation** | **Universal response** | **Heterogeneous response** |


## Individual Variability: Universal F0 Elevation vs Heterogeneous NNE Responses

### Contrasting Patterns of Individual Differences

A critical finding emerged from variance partitioning: stress-induced F0 elevation was 
remarkably consistent across individuals (τ₂/β₁ = 0.33; 95% PI: 1.2–5.4 Hz, all positive), 
whereas NNE responses showed pronounced heterogeneity (τ₂/β₁ = 1.27; 95% PI: -2.3 to +1.0 dB, 
crossing zero). This fourfold difference in relative variability indicates fundamentally 
different processes underlying these two dimensions of vocal stress response.

Most strikingly, while 100% of individuals exhibited stress-related pitch elevation, our 
model estimates that approximately 22% showed stress-related increases in glottal noise—
directly opposite to the population-level trend toward reduced noise. This is not merely 
statistical noise: individuals showing "paradoxical" NNE increases may represent a distinct 
subgroup employing qualitatively different phonatory strategies under stress.

### Theoretical Implications

**Why is F0 universal but NNE heterogeneous?**

The universality of F0 elevation likely reflects the obligatory nature of autonomic stress 
responses. Sympathetic activation increases laryngeal muscle tension and subglottal pressure 
via relatively automatic, phylogenetically conserved pathways that operate similarly across 
individuals. While the magnitude varies (presumably reflecting trait differences in 
autonomic reactivity), the direction is invariant.

In contrast, NNE—indexing the balance between harmonic and noise energy in the glottal 
signal—depends on fine-grained adjustments in vocal fold contact patterns, adduction force, 
and phonatory mode. These adjustments likely reflect a combination of:

1. **Compensatory vocal strategies**: Some speakers may habitually adopt pressed phonation 
   under stress (reducing noise), while others maintain modal phonation or even shift toward 
   breathy voice (increasing noise)

2. **Baseline vocal habits**: Individuals with habitually pressed or tense voice quality may 
   show ceiling effects, limiting further noise reduction

3. **Attentional modulation**: Stress-induced hypervigilance might promote conscious vocal 
   monitoring in some individuals (enhancing control, reducing noise) but distraction in 
   others (degrading control, increasing noise)

4. **Physiological constraints**: Anatomical differences in laryngeal structure or chronic 
   vocal tension may constrain the range of phonatory adjustments available under stress

### Practical and Clinical Implications

These findings have important implications for applied stress detection:

**F0 as a population-level biomarker**: The consistency of stress-related F0 elevation across 
all individuals—despite variation in magnitude—suggests F0 could serve as a relatively 
robust, calibration-free indicator of acute stress in real-world applications.

**NNE requires individual calibration**: The 22% of individuals showing opposite effects means 
NNE cannot be interpreted without individual baseline data or additional contextual information. 
This may explain why personality trait moderators (examined separately) showed stronger effects 
for F0 than NNE: traits may predict magnitude of universal responses more reliably than 
direction of heterogeneous responses.

**Multi-parameter approaches essential**: The dissociation underscores that vocal stress 
responses are multidimensional. Relying solely on NNE (or any single parameter) risks 
misclassifying the ~22% of individuals whose responses deviate from population trends. 
Combining F0 (universal) with NNE (heterogeneous but informative about individual strategies) 
provides complementary information about both general arousal state and individual-specific 
regulatory dynamics.

### Unanswered Questions

Why do ~22% of individuals show increased glottal noise under stress? Potential explanations 
include:

- **Vocal fatigue**: If stress-induced tension leads to incomplete glottal closure in some 
  individuals
- **Breathy phonation as anxiety marker**: Some individuals may adopt breathy voice as a 
  submissive or affiliative signal under evaluative stress
- **Measurement confounds**: Increased respiratory rate under stress might introduce 
  turbulent noise that NNE captures but that reflects respiration rather than phonation

Future work should:
1. Examine whether the 22% showing paradoxical NNE increases differ systematically in 
   personality, vocal training, or baseline voice quality
2. Use laryngeal imaging or EGG to directly observe glottal dynamics in high vs low NNE 
   responders
3. Test whether individual NNE response patterns are stable across stressors or represent 
   state-dependent variation
   
\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
